---
title: "Statistical Inference Project"
author: "Alfonso Gonzalez Damian"
subtitle: Exponential distribution vs CLT
output: html_document
---

### Exploring dataset
* Exploring the dataset with summary, dim and plot.
* We can see two clear groups in tooth growth by supp and three groups by dose
```{r}
data(ToothGrowth)
library(reshape2)
summary(ToothGrowth)
dim(ToothGrowth)
n <- length(ToothGrowth$len)
par(mfrow =c(1,2))
plot(ToothGrowth$len ~ ToothGrowth$supp, type="l", xlab="Supp", ylab="Tooth Growth")
plot(ToothGrowth$len ~ as.factor(ToothGrowth$dose), xlab="Dose", ylab="Tooth Growth")
```
* We can test if this sample is a good estimator of the real population differences in toothgrowht

### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
* Using the t.test in R for the hypothesis test that there is a difference greater than 0:

```{r}
results <- t.test(len ~ supp, paired = FALSE, alternative="g", data= ToothGrowth)
results
```
* As we can see in the results the p-value < 0.05, so we can accept the alternative hypothesis

### Conclusions and the assumptions needed
We can conclude that the confidence interval for the two groups (OC and VC) mean difference in toot growth with a 95% of confidence includes the true difference between them.